---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, eval=FALSE}
library(dplyr)


DF <- ChickWeight %>%
  filter(Diet %in% c(3,4))

DF

DF %>%
  group_by(Diet) %>%
  summarise(mean(weight), var(weight), n())



t.test(weight~Diet, data=DF)
```


if p < alpha (defaul to 0.05), reject the null hypo
reject null, or fail to reject. no confirming
conclusion always in terms of alt hypo

type 1 error: reject h0 when h0 is true
type 2 error: fail to reject h0 when h0 is false


#GGPLOT ----------------------------------------------

* boxplot
* bargraph
* jitter/point



#BINOM --------------------------------------------------
#### * Generate 100 occurrences of flipping 10 coins, each with 30% probability
  rbinom(100, 10, .3)
  rbinom(# of repeats, # of flips, % chance)

#### * Calculate the probability that 2 are heads using dbinom
    dbinom(2, 10, .3)  
    dbinom(# equal to heads, # of flips, % chance)  
    
      equivalent to:
       Confirm your answer with a simulation using rbinom
      mean(rbinom(10000, 10, .3) == 2)
  
#### * Calculate the probability that at least five coins are heads
  1 - pbinom(4, 10, .3)
  pbinom(# <= heads, # of flips, % chance)
         
      equivalent to:
       Confirm your answer with a simulation of 10,000 trials
      mean(rbinom(10000, 10, .3) >= 5)
      
#### * Simulate 100,000 flips of a coin with a 40% chance of heads
  A <- rbinom(100000, 1, .4) 

   Simulate 100,000 flips of a coin with a 20% chance of heads
  B <- rbinom(100000, 1, .2)

   Estimate the probability both A and B are heads
  mean(A & B)      
  
  
  
  
      
#PVALUE ---------------------------------------------------
    the probability that you observe a value as extreme as ___ based on the null hypothesis being true.
    if p < alpha, reject null.
    if p > alpha, fail to reject null
    
    if alt hypotheseis is 2 sided (!=), must multiply p value by 2
    
#RANDOM NOTES-----------------------------------------------

```{r}
#soda bootstrapping/confidence interval notes from thursday
sims <- 10000
bs <- numeric(sims)
for (i in 1:sims)
{
  west_coast <- mean(sample(0:1, size=26, replace=TRUE, prob=c(7/26, 19/26)))
  east_coast <- mean(sample(0:1, size=34, replace=TRUE, prob=c(6/34, 28/34)))
  
  bs[i] <- west_coast-east_coast  
}

hist(bs)
quantile(bs, prob=c(.025, .975))
#creates a bootstrapped 95% confidence interval 

```


fundamental question of inference -- how does what you actually observe compare to what would happen if the null
  hypothesis were true, and you repeat the process many times?
  
  probability of A intersect prob B has to = prob A * prob B if A and B are independent
  
```{r}
time <- c(30,25,20,18,21,22)
obs_stat <- mean(time[1:3]) - mean(time[4:6])
obs_stat

#as fraction
MASS::fractions(obs_stat)

sims <- 10^4 - 1

#make answer vector
ans <- numeric(sims)

for(i in 1:sims)
{
  #returns 3 random vals between 1 and 6 without repeating
  index <- sample(6, 3, replace = FALSE)
  
  #puts 3 random vals from time into ans
    ans[i] <- mean(time[index]) - mean(time[-index])
}

pvalue <- (sum(ans >= obs_stat) + 1) / (sims + 1)
pvalue
```
if fail to reject null: the evidence fails to suggest there is a difference between the drug and control groups
if reject null: the evidence suggests the mean time for a mouse on drugs is greater 
  
```{r}
ans <- combn(time, 3)
ans
avg_ans <- numeric(20)
for(i in 1:20)
{
   avg_ans[i] <- mean(ans[,i])
}
avg_ans
```
  
```{r, eval=FALSE}
# loop through matrix
for (i in 1:nrow(matrix)) {
  for (j in 1:ncol(matrix)) {
    print(paste("On row", i, "and column", j, "the board contains", matrix[i, j]))
  }
}
```
  
  