---
title: "Chapter Three Homework Solutions"
author: "michael walker, francis boadu"
date: '`r format(Sys.time(), "%b %d, %Y")`'
output: bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, fig.align = "center", warning = FALSE, message = FALSE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(oilabs)
```

1. In the Flight Delays Case Study in Section 1.1,

    a. The data contain flight delays for two airlines, American Airlines and United Airlines.  Conduct a two-sided permutation test to see if the mean delay times between the two carriers are statistically significant.
    
    b. The flight delays occured in May and June of 2009.  Conduct a two-sided permutation test to see if the difference in mean delay times between the 2 months is statistically significant.
    
```{r}
FD <- read.csv("http://www1.appstate.edu/~arnholta/Data/FlightDelays.csv")
# a. Your code here
FD <- FD %>%
  group_by(Carrier)

obs_stat <- mean(FD$Delay[1:1123]) - mean(FD$Delay[1124:4029])
obs_stat

sims <- 10^4 - 1

#make answer vector
ans <- numeric(sims)

for(i in 1:sims)
{
  index <- sample(nrow(FD), 1123, replace = FALSE)
  
    ans[i] <- mean(FD$Delay[index]) - mean(FD$Delay[-index])
}

pvalue <- (sum(ans >= obs_stat) + 1) / (sims + 1)
pvalue



``` 



```{r}
# b. Your code here
FD <- read.csv("http://www1.appstate.edu/~arnholta/Data/FlightDelays.csv")

FD <- FD %>%
  group_by(Month)

obs_stat <- mean(FD$Delay[1:1999]) - mean(FD$Delay[2000:4029])
obs_stat

sims <- 10^4 - 1

ans <- numeric(sims)

for(i in 1:sims)
{
  index <- sample(nrow(FD), 1999, replace = FALSE)
  
    ans[i] <- mean(FD$Delay[index]) - mean(FD$Delay[-index])
}

pvalue <- (sum(ans >= obs_stat) + 1) / (sims + 1)
pvalue

```

    
2. In the Flight Delays Case Study in Section 1.1, the data contain flight delays for two airlines, American Airlines and United Airlines.

    a. Compute the proportion of times that each carrier's flights was delayed more than 20 minutes.  Conduct a two-sided test to see if the difference in these proportions is statistically significant.
    
    b. Compute the variance in the flight delay lengths for each carrier.  Conduct a test to see if the variance for United Airlines is greater than that of American Airlines.
    
```{r}
# a. Your code here
FD <- read.csv("http://www1.appstate.edu/~arnholta/Data/FlightDelays.csv")

FD <- FD %>%
  group_by(Carrier)
FD %>%  summarize(delay_prop = mean(Delay > 20))


obs_stat <- mean(FD$Delay[1:1123] > 20) - mean(FD$Delay[1124:4029] > 20)
obs_stat

sims <- 10^4 - 1

#make answer vector
ans <- numeric(sims)

for(i in 1:sims)
{
  index <- sample(nrow(FD), 1123, replace = FALSE)
  
    ans[i] <- mean(FD$Delay[index] > 20) - mean(FD$Delay[-index] > 20)
}

pvalue <- (sum(ans >= obs_stat) + 1) / (sims + 1)
pvalue


```


```{r}
# b. Your code here
FD <- read.csv("http://www1.appstate.edu/~arnholta/Data/FlightDelays.csv")

FD %>%
  group_by(Carrier) %>%
  summarize(Variance = var(Delay))

var.test(FD$Delay[1:1123], FD$Delay[1124:4029], alternative = "two.sided")
```

```{r}
FD %>%  summarize(delay_prop = var(Delay))


obs_stat <- var(FD$Delay[1:1123]) - var(FD$Delay[1124:4029])
obs_stat

sims <- 10^4 - 1

#make answer vector
ans <- numeric(sims)

for(i in 1:sims)
{
  index <- sample(nrow(FD), 1123, replace = FALSE)
  
    ans[i] <- var(FD$Delay[index]) - var(FD$Delay[-index])
}

pvalue <- (sum(ans >= obs_stat) + 1) / (sims + 1)
pvalue

```

    
    
3. In the Flight Delays Case Study in Section 1.1, repeat Exercise 3 part (a) using three test statistics: (i) the mean of the United Airlines delay times, (ii) the sum of the United Airlines delay times, and (iii) the difference in the means, and compare the P-values. Make sure all three test statistics are computed within the same `for` loop.

```{r}
# Your code here
FD <- read.csv("http://www1.appstate.edu/~arnholta/Data/FlightDelays.csv")

FD <- FD %>%
  filter(Carrier == "UA")

UA_mean <- FD %>% mean(Delay)
UA_sum <- sum(FD$Delay)
mean_diff <- mean(FD$Delay[1:1123]) - mean(FD$Delay[1124:4029])
  
  

sims <- 10^4 - 1

#make answer vector
ans1 <- numeric(sims)

for(i in 1:sims)
{
  index <- sample(nrow(FD), 1123, replace = FALSE)
  
    ans1[i] <- mean(FD$Delay[index]) - mean(FD$Delay[-index])
}

pvalue1 <- (sum(ans >= UA_mean) + 1) / (sims + 1)
pvalue2 <- (sum(ans >= UA_sum) + 1) / (sims + 1)
pvalue3 <- (sum(ans >= mean_diff) + 1) / (sims + 1)


pvalue1
pvalue2
pvalue3
```


4. In the Flight Delays Case Study in Section 1.1,
    
    a. Find the 25% trimmed mean of the delay times for United Airlines and American Airlines.
    
    b. Conduct a two-sided test to see if the difference in trimmed means is statistically significant.
    
```{r}
# Your code here
FD <- read.csv("http://www1.appstate.edu/~arnholta/Data/FlightDelays.csv")

FD <- FD %>%
  group_by(Carrier) 
FD %>%  summarize(trimmed_mean = mean(Delay, .25))


obs_stat <- mean(FD$Delay[1:1123], .25) - mean(FD$Delay[1124:4029], .25)
obs_stat

sims <- 10^4 - 1

#make answer vector
ans <- numeric(sims)

for(i in 1:sims)
{
  index <- sample(nrow(FD), 1123, replace = FALSE)
  
    ans[i] <- mean(FD$Delay[index], .25) - mean(FD$Delay[-index], .25)
}

pvalue <- (sum(ans >= obs_stat) + 1) / (sims + 1)
pvalue

```
    
    
5. In the Flight Delays Case Study in Section 1.1,

    a. Compute the proportion of times the flights in May and in June were delayed more than 20 min, and conduct a two-sided test of whether the difference between months is statistically significant.
    
    b. Compute the variance of the flight delay times in May and June and then conduct a two-sided test of whether the ratio of variances is statistically significantly different from 1.
    
```{r}
# a. Your code here
FD <- read.csv("http://www1.appstate.edu/~arnholta/Data/FlightDelays.csv")

FD <- FD %>%
  group_by(Month) 
FD %>%  summarize(mean(Delay > 20))


obs_stat <- mean(FD$Delay[1:1999] > 20) - mean(FD$Delay[2000:4029] > 20)
obs_stat

sims <- 10^4 - 1

#make answer vector
ans <- numeric(sims)

for(i in 1:sims)
{
  index <- sample(nrow(FD), 1999, replace = FALSE)
  
    ans[i] <- mean(FD$Delay[index] > 20) - mean(FD$Delay[-index] > 20)
}

pvalue <- (sum(ans >= obs_stat) + 1) / (sims + 1)
pvalue

```

```{r}
# b. Your code here
FD <- read.csv("http://www1.appstate.edu/~arnholta/Data/FlightDelays.csv")

FD %>%
  summarize(delay_var = var(Delay))

obs_stat <- var(FD$Delay[1:1999]) - var(FD$Delay[2000:4029])
obs_stat

sims <- 10^4 - 1

#make answer vector
ans <- numeric(sims)

for(i in 1:sims)
{
  index <- sample(nrow(FD), 1999, replace = FALSE)
  
    ans[i] <- var(FD$Delay[index]) - mean(FD$Delay[-index])
}

pvalue <- (sum(ans >= obs_stat) + 1) / (sims + 1)
pvalue

```

6. Research at the University of Nebraska conducted a study to investigate sex differences in dieting trends among a group of Midwestern college students (Davy et al. (2006)). Students were recruited from an introductory nutrition course during one term. Below are data from one question asked to 286 participants.

    a. Write down the appropriate hypothesis to test to see if there is a relationship between gender and diet and then carry out the test.
   $H_0$: Proportion of women dieting is equal to the proportion of men dieting
   $H_A$ Proportion of women dieting is not equal to the proportion of men dieting
   
    b. Can the resluts be generalized to a population? Explain.
    Yes; bootstrapping allows approximating a population from a sample
    
```{r, echo = FALSE}
DT <- matrix(c(35, 146, 8, 97),nrow=2, byrow=TRUE)
dimnames(DT) <- list(Gender =c("Women", "Men"), LowFatDiet = c("Yes", "No"))
DT
```

```{r}
# Your code here
chisq.test(DT, correct = FALSE)
```

7. A national polling company conducted a survey in 2001 asking a randomly selected group of Americans of 18 years of age or older whether they supported limited use of marijuana for medicinal purposes. Here is a summary of the data:

    Write down the appropriate hypothesis to test whether there is a relationship between age and support for medicinal marijuana and carry out the test.
    $H_0$: Proportion of advocates is equal across age groups
    $H_A$: Proportion of advocates is not equal across age groups
  
```{r, echo = FALSE}
MA <- matrix(c(172, 52, 313, 103, 258, 119), nrow = 3, byrow = TRUE)
dimnames(MA) <- list(Age = c("18-29 years old", "30-49 years old", "50 years or older"), Support = c("For", "Against"))
MAT <- as.table(MA)
MADF <- as.data.frame(MAT)
DF <- as.tbl(vcdExtra::expand.dft(MADF))
T1 <- xtabs(~Age + Support, data = DF)
T1
```  



```{r}
# Your code here
chisq.test(T1, correct = FALSE)
```


8. Two students went to a local supermarket and collected data on cereals; they classified by their target consumer (children versus adults) and the placement of the cereal on the shelf (bottom, middle, and top). The data are given in `Cereals`.

    a. Create a table to summarize the relationship between age of target consumer and shelf location.
    
    b. Conduct a chi-square test using R's `chisq.test` command.
    
    c. `R` returns a warning message. Compute the expected counts for each cell to see why.
    
    d. Conduct a permutation test for independence.
    
```{r, message = TRUE}
Cereals <- read.csv("http://www1.appstate.edu/~arnholta/Data/Cereals.csv")
# Your code here
TF <- xtabs(~Age + Shelf, data = Cereals)
TF
chisq.test(TF, correct = FALSE)

obs_stat <- chisq.test(TF, FALSE)$statistic
obs_stat

sims <- 10^4 - 1

#make answer vector
ans <- numeric(sims)

for(i in 1:sims)
{
  index <- xtabs(~sample(Age) + Shelf, data = Cereals)
  
    ans[i] <- chisq.test(index)$statistic
}

pvalue <- (sum(ans >= obs_stat) + 1) / (sims + 1)
pvalue
```


9. From GSS 2002 Case Study in Section 1.6,
    
    a. Create a table to summarize the relationship between gender and the person's choice for president in the 2000 election.
    
    b. Test to see if a person's choice for president in the 2000 election is independent of gender (use chisq.test in `R`).
    
    c. Repeat the test but use the permutation test for independence. Does your conclusion change? (Be sure to remove observations with missing values)
    
```{r}
GSS2002 <- read.csv("http://www1.appstate.edu/~arnholta/Data/GSS2002.csv")
# Your code here

TF <- xtabs(~Gender + Pres00, data = GSS2002) 

chisq.test(TF)

chisq.test(TF, correct = FALSE)

obs_stat <- chisq.test(TF, FALSE)$statistic
obs_stat

sims <- 10^4 - 1

#make answer vector
ans <- numeric(sims)

for(i in 1:sims)
{
  index <- xtabs(~sample(Gender) + Pres00, data = GSS2002)
  
    ans[i] <- chisq.test(index)$statistic
}

pvalue <- (sum(ans >= obs_stat) + 1) / (sims + 1)
pvalue

```
    

10. From GSS 2002 Case Study in Section 1.6,

    a. Create a table to summarize the relationship bewteen gender and the person's general level of happiness (`Happy`).
    
    b. Conduct a permutation test to see if gender and level of happiness are independent (Be sure to remove the observations with missing values).

```{r}
# Your code here
TF <- xtabs(~Gender + Happy, data = GSS2002)

chisq.test(TF, correct = FALSE)

obs_stat <- chisq.test(TF, FALSE)$statistic
obs_stat

sims <- 10^4 - 1

#make answer vector
ans <- numeric(sims)

for(i in 1:sims)
{
  index <- xtabs(~sample(Gender) + Happy, data = GSS2002)
  
    ans[i] <- chisq.test(index)$statistic
}

pvalue <- (sum(ans >= obs_stat) + 1) / (sims + 1)
pvalue

```

    
11. From GSS 2002 Case Study in Section 1.6,

    a. Create a table to summarize the relationship between support for gun laws (`GunLaw`) and views on government spending on the military (`SpendMilitary`).
    
    b. Conduct a permutation test to see if support for gun laws and views on government spending on the military are independent (Be sure to remove observations with missing values).

    
```{r}
# Your code here
TF <- xtabs(~GunLaw + SpendMilitary, data = GSS2002)

chisq.test(TF, correct = FALSE)

obs_stat <- chisq.test(TF, FALSE)$statistic
obs_stat

sims <- 10^4 - 1

#make answer vector
ans <- numeric(sims)

for(i in 1:sims)
{
  index <- xtabs(~sample(GunLaw) + SpendMilitary, data = GSS2002)
  
    ans[i] <- chisq.test(index)$statistic
}

pvalue <- (sum(ans >= obs_stat) + 1) / (sims + 1)
pvalue

```
    
    