---
title: "R Markdown Practice"
author: "Michael Walker"
date: "Sep, 13, 2017"
output: 
  pdf_document: 
    fig_height: 1.5
    fig_width: 2.5
---


Directions: Recreate this document using R Markdown. Make sure that you use inline R to report your answers. Your document should look like this document when it is knitted including the directions but have your name in place of   the current *Your Name*. Please print (before class) and turn in both the `*.Rmd file` and the knitted `*.pdf` file stapled to the back of your `*.Rmd` file at the start of class 9/14/17. Name your file `firstname_lastname.Rmd` (mine would be `alan_arnholt.Rmd`). Use global options to set the height and width of your figures to 1.5 and 2.5 inches, respectively.


##Some Code


```{r graph_stuff, include=FALSE}
set.seed(31)
x <- rnorm(1000, 100, 10)
DF <- data.frame(x = x)
library(ggplot2)
ggplot(data = DF, aes(x = x)) +
  geom_histogram(binwidth = 2, fill = "pink", color = "black") +
  theme_bw() + theme(text=element_text(family="Times", size=12))
```
```{r ref.label='graph_stuff', eval=FALSE}
```

The mean of the graph shown below is $\bar{x} =$ `r round(mean(DF$x), 2)`. The standard deviation of the graph below is $s =$`r round(sd(DF$x), 2)`. Make sure your answers update properly and are rounded to two decimal places when the value passed to `set.seed()` changes.

```{r echo=TRUE, comment=NA}
summary(DF$x)
```

The third quartile, $Q_3$, is `r round(summary(DF$x)[5], 1)`

##A Graph

```{r ref.label='graph_stuff', echo=FALSE, fig.align='center'}

```
  
##Additional Resources

* http://rmarkdown.rstudio.com/
* Cheat Sheets