---
title: "R Markdown Practice"
author: Michael Walker
date: "10/3/2017"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, fig.align = "center", comment=NA)
library(dplyr)
library(ggplot2)
library(PASWR2)
```


#Is there a relationship between area and price?
To see the relatioship between `totalprice` and `area` see Figure \@ref(fig:g1).
```{r, label="g1", fig.cap="scatterplot"}
ggplot(data = VIT2005, aes(x=area, y=totalprice)) + geom_point(color="lightblue") + theme_bw() +geom_smooth(method = "lm", se=FALSE)
```
* Direction
* Form
* Strength


##Correlation
$-1 \leq r \leq 1$

##Least Squares
The equation for a simple linear regression is given in \@ref(eq:ls)

\begin{equation}
\hat{y}_i = b_0 + b_1 \cdot x_i
(\#eq:ls)
\end{equation}

##The lm function
```{r, eval=FALSE}
model <- lm(response ~predictor, data=DataFrame)
```

```{r}
model <- lm(totalprice~area, data=VIT2005)
#area 2705 = slope of line
model
summary(model)
coefficients(summary(model))
```

The price of the apartment goes up by `r round(coefficients(summary(model))[2,1],2)`
euros for each $\text{m}^2$
unit increase